Data Preparation

### 2023-03-30 x_KATL.qmd
building study data loop

---
title: "Untitled"
format: html
---

@@

@@

study_trjs <- katl_trjs |> filter(ICAO24 %in% icaos_seen$ICAO24) |> 
  coerce_unixepoch_to_datetime() |> 
  identify_trajectory_legs() |> 
  mutate(TID = paste(ICAO24, LEG, sep = "-"))

@@

study_trjs <- study_trjs |> filter(TID %in% heuristic_arrs$TID)  
```

>>>>

---
title: "Spacing Deviation at US airports"
format: html
---

```{r}
library(tidyverse)

files_to_sources = list.files(path = "./R", full.names = TRUE)
purrr::walk(files_to_sources, .f = source)
```

@@

katl_200NM <- trrrj::polygon_at_distance(c(arp_katl$LON, arp_katl$LAT), d = 200) |> 
  sf::st_as_sf() |> mutate(REF = "KATL-200NM")

@@

study_trjs <- katl_trjs |> filter(ICAO24 %in% icaos_seen$ICAO24) |> 
  coerce_unixepoch_to_datetime() |> 
  identify_trajectory_legs() |> 
 # fix_invalid_fltid() |> 
  mutate(TID = paste(ICAO24, LEG, sep = "-"))

@@

study_trjs <- study_trjs |> filter(TID %in% heuristic_arrs$TID)

arrivals <- study_trjs |> 
  assign_landing_runway(
      .arp = arp_katl
    , .rwy_centerline_ls = katl_ctrl_ls 
    , .upper_bound = 10
    , .group_var = TID
    ) |> 
  filter(N > 5)    # for 10 sec update crossers have typically low number of hits
```
```{r}
itr <- arrivals[1,] |> select(UID = TID, UID_REF = REF)

make_study_arrival_trajectories <- function(.itr, .trjs, .rwys){
  message("\n", itr$UID)
  app_trj <- .trjs |> dplyr::filter(UID == itr$UID)
  app_thr <- .rwys |> dplyr::filter(REF == itr$UID_REF)

  # estimate threshold (aka landing time)
  thr_row <- estimate_threshold_time(app_trj, app_thr) |> 
    select(UID, ICAO24, FLTID, TIME, LAT, LON ) |> 
    mutate(MST = "THR", MST_LABEL = itr$UID_REF)

  # append threshold row and trim trajectory
  app_trj <- app_trj |> bind_rows(thr_row) |> 
    select(UID, ICAO24, FLTID, TIME, LAT, LON, ALT_B, MST, MST_LABEL) |> 
    arrange(TIME) |> 
    rowid_to_column("SEQ_ID")

  # add "distance-to-go" and "time-to-fly" 
  app_trj <- app_trj |> add_flight_distance()
  return(app_trj)
}

bingo <- itr |> make_study_arrival_trajectories(study_trjs |> rename(UID = TID), rwys_katl)
```

run for many, function re-written to add RWY to trjs

```{r}
study_payload <- study_trjs |> 
  select(TID, ICAO24, FLTID, TIME, LAT, LON, ALT_B) |> 
  left_join(arrivals |> select(TID, RWY = REF), by = join_by(TID)) |> 
  rename(UID = TID)

rqq <- study_payload |> 
  mutate(LUID = UID) |> 
  group_by(LUID) |> 
  group_map(.f = ~ safer_make_study_arrival_trajectories2( ., .rwys = rwys_katl))

# check for errors
names(rqq) <- seq_along(rqq)
keep(rqq, ~ ! is.data.frame(.x))
```

```{r}
next_payload <- keep(rqq, ~is.data.frame(.x)) |> bind_rows()
```

### 2023-03-30 x_bra-check.qmd
building study data loop

```{}
# load helper scripts to download and store OSN data
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```
```{}
#| message: false
pth_to_dumps <- "../__DATA/OSN-MondayDumps"
ptn <- "2022-05-09"

@@

```{r}
katl_arp   <- c(-84.4278640,33.6366996)
katl_200NM <- trrrj::polygon_at_distance(katl_arp, d = 205) |> sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "KATL-205")

>>>>

```{}
# load helper scripts to download and store OSN data
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```
```{r}
#| message: false
pth_to_dumps <- "../__DATA/OSN-MondayDumps"
ptn <- "2022-05-09"

@@

```{r}
katl_arp   <- c(-84.4278640,33.6366996)
arp_katl   <- tribble(
  ~LAT,   ~LON,~REF,~ELEV
  ,33.6, -84.4,"ARP", 1026
)

# SPOILER =========
# trrrj polygon at uses "ireland crs"! 
# katl_200NM <- trrrj::polygon_at_distance(katl_arp, d = 205) |> sf::st_as_sf() |> 
#   rename(geometry = x) |> 
#   mutate(NAME = "KATL-205")
katl_200NM <- arp_katl |> select(LAT, LON) |> 
  cast_latlon_to_pts() |> sf::st_buffer(205 * 1852) |> 
  mutate(REF = "KATL-205NM")

### 2023-03-30 x_data-crunching.qmd
building study data loop

```{r}
icao24_in_polygon <- function(.ds, .poly, .id = ICAO24){
  # determine bounding box for polygon
  this_box <- sf::st_bbox(.poly)
  # pre-filter with dplyr - cut bounding box
  icao24s <- .ds |> 
    dplyr::filter(
       LAT > this_box$ymin, LAT < this_box$ymax
      ,LON > this_box$xmin, LON < this_box$xmax, 
    )
  # get ids seen
  icao24s <- icao24s |> 
    dplyr::summarise(
      N = n(), N_ID = length(unique(FLTID))
      , MIN_TIME = min(TIME,na.rm = TRUE)
      , MIN_TIME_ALT = ALT_B[which(TIME == MIN_TIME)]
      , MAX_TIME = max(TIME, na.rm = TRUE)
      , MAX_TIME_ALT = ALT_B[which(TIME == MAX_TIME)]
      , MIN_ALT = min(ALT_B, na.rm = TRUE)
      , MED_ALT = median(ALT_B, na.rm = TRUE)
      , MAX_ALT = max(ALT_B, na.rm = TRUE)
      , .by = {{.id}})
}
```

>>>>

```{r}
source("~/RProjects/paper-2023-ICNS/R/trj_icao24_in_polygon.R")
# MOVED TO R
# ==================
# icao24_in_polygon <- function(.ds, .poly, .id = ICAO24){
#   # determine bounding box for polygon
#   this_box <- sf::st_bbox(.poly)
#   # pre-filter with dplyr - cut bounding box
#   icao24s <- .ds |> 
#     dplyr::filter(
#        LAT > this_box$ymin, LAT < this_box$ymax
#       ,LON > this_box$xmin, LON < this_box$xmax, 
#     )
#   # get ids seen
#   icao24s <- icao24s |> 
#     dplyr::summarise(
#       N = n(), N_ID = length(unique(FLTID))
#       , MIN_TIME = min(TIME,na.rm = TRUE)
#       , MIN_TIME_ALT = ALT_B[which(TIME == MIN_TIME)]
#       , MAX_TIME = max(TIME, na.rm = TRUE)
#       , MAX_TIME_ALT = ALT_B[which(TIME == MAX_TIME)]
#       , MIN_ALT = min(ALT_B, na.rm = TRUE)
#       , MED_ALT = median(ALT_B, na.rm = TRUE)
#       , MAX_ALT = max(ALT_B, na.rm = TRUE)
#       , .by = {{.id}})
# }
```
